---
title: "Raw data"
author: "Javier Porras"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
editor: visual
bibliography: "C:/Users/javier.porras/Documents/arkaute_/code/experiment_references.bib"
---


# Data treatment for Arkaute's experiment of recovery dynamics of herbaceous community and climate change

## Context

In the experiment carried out in Arkaute (Alava), the 1 m$^2$.... \[insert all information about biomass limitations, chosing non-destructive methods and bibliography\]. @pottier2017,

## Checking 

First thing that we can do, it's checking te distribution of the morphological parameters measured Cb, Cm, Dm, Db. 

```{r sdfsdf}
#| echo: false
#| warning: false


rm(list = ls(all.names = TRUE))
pacman::p_load(dplyr,reshape2,tidyverse, lubridate, ggplot2, ggpubr, rpivotTable, ggrepel, broom)

setwd("C:/Users/javier.porras/Documents/arkaute_")


theme_set(theme_bw() +
            theme(
              legend.position = "right",
              panel.grid = element_blank(),
              strip.background = element_blank(),
              strip.text = element_text(face = "bold"),
              text = element_text(size = 11)))




# OPENING DATA #####

flora_raw <- read.csv("data/flora_db_raw.csv") # Opening and transforming data(opening_floradata.R) ####

flora_raw <- flora_raw %>%
  mutate(across(where(is.character), as.factor))

flora_raw$plot <- factor(flora_raw$plot)


flora_raw$sampling <- factor(flora_raw$sampling, levels = sort(unique(flora_raw$sampling)))
flora_raw$treatment <- factor(flora_raw$treatment, levels =  c("c", "w", "p", "wp") )
flora_raw$plot <- factor(flora_raw$plot, levels = sort(unique(flora_raw$plot)))

flora_raw <- select(flora_raw, -date, -category, -OBS, -ni.agrupado, -familia, -species_old_1, -species_old_2)
```


```{r}
#| echo: false
#| warning: false
#| fig-cap: "Fig.1: Boxplots and histograms for morphological parameters"

par(mfrow = c(2,2))
boxplot(flora_raw$Dm, main = "Dm")
boxplot(flora_raw$Cm, main = "Cm")
boxplot(flora_raw$Db, main = "Db")
boxplot(flora_raw$Cb, main = "Cb")
```


```{r }

par(mfrow = c(2,2))
hist(flora_raw$Dm, main = "Dm", breaks = 50)
hist(flora_raw$Cm, main = "Cm", breaks = 50)
hist(flora_raw$Db, main = "Db", breaks = 50)
hist(flora_raw$Cb, main = "Cb", breaks = 50)
```


```{r }
par(mfrow = c(2,1))
boxplot(flora_raw$abundance, main = "Abundance")
hist(flora_raw$abundance, main = "Abundance", breaks = 50)

```







